---
title: "Create database from data frame `mtcars`"
subtitle: "Using base R"
author: "Alex"
date: "`r format(Sys.time(), '%d %B %Y, %H:%M:%S')`"
output: 
  html_document:
    df_print: paged
    toc: true
    number_sections: false
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(RSQLite)
```

Print data frame overview.

```{r}
head(mtcars)
```

Load helper functions.

```{r}
source(file = "helper.R")
```

Create augmented data frame.

```{r}
z <- within(mtcars, {
  make <- rownames(mtcars)
  cyl_id <- as.intfac(cyl)
  vs_id <- as.intfac(vs)
  am_id <- as.intfac(am)
  gear_id <- as.intfac(gear)
  carb_id <- as.intfac(carb)
})

z <- add_id_col(z)

# Reorder columns
z <- z[, c(names(mtcars), "make", "id", "cyl_id", "vs_id", "am_id", "gear_id", "carb_id")]

rownames(z) <- seq_len(nrow(z))
head(z)
```

Create table of cars specs (mpg, disp, hp, drat, wt, qsec).

```{r}
specs_table <- z[, c("id", "mpg", "disp", "hp", "drat", "wt", "qsec")]
head(specs_table)
```

Create table of cyl.

```{r}
cyl_table <- z[, c("cyl", "cyl_id")] |>
  unique() |>
  (\(df) df[order(df$cyl_id), ])() |>
  (\(df) { df$cyl <- as.integer(df$cyl); df })() |>
  (\(df) { rownames(df) <- NULL; df })()
cyl_table
```

Create table of vs.

```{r}
vs_table <- z[, c("vs", "vs_id")] |>
  unique() |>
  (\(df) df[order(df$vs_id), ])() |>
  (\(df) { df$vs <- as.integer(df$vs); df })() |>
  (\(df) { rownames(df) <- NULL; df })()
vs_table
```

Create table of am.

```{r}
am_table <- z[, c("am", "am_id")] |>
  unique() |>
  (\(df) df[order(df$am_id), ])() |>
  (\(df) { df$am <- as.integer(df$am); df })() |>
  (\(df) { rownames(df) <- NULL; df })()
am_table
```

Table of gear.

```{r}
gear_table <- z[, c("gear", "gear_id")] |>
  unique() |>
  (\(df) df[order(df$gear_id), ])() |>
  (\(df) { df$gear <- as.integer(df$gear); df })() |>
  (\(df) { rownames(df) <- NULL; df })()
gear_table
```

Table of carb.

```{r}
carb_table <- z[, c("carb", "carb_id")] |>
  unique() |>
  (\(df) df[order(df$carb_id), ])() |>
  (\(df) { df$carb <- as.integer(df$carb); df })() |>
  (\(df) { rownames(df) <- NULL; df })()
carb_table
```

Table of cars.

```{r}
cars_table <- z[, c("id", "make", "cyl_id", "vs_id", "am_id", "gear_id", "carb_id")] |>
  unique() |>
  (\(df) df[order(df$id), ])() |>
  (\(df) { rownames(df) <- NULL; df })()
cars_table
```

Rejoin tables to recreate original data frame.

```{r}
mtcars_merged <- cars_table |> 
  merge(cyl_table, by = "cyl_id", all.y = TRUE) |> 
  merge(vs_table, by = "vs_id", all.y = TRUE) |> 
  merge(am_table, by = "am_id", all.y = TRUE) |> 
  merge(gear_table, by = "gear_id", all.y = TRUE) |> 
  merge(carb_table, by = "carb_id", all.y = TRUE) |> 
  (\(df) df[c("id", "make", "cyl", "vs", "am", "gear", "carb")])() |> 
  merge(specs_table, by = "id", all.y = TRUE)
head(mtcars_merged)
```

Restore row names.

```{r}
rownames(mtcars_merged) <- z$make
```

Reorder columns and fix data types.

```{r}
cols_to_num <- c("cyl", "vs", "am", "gear", "carb")
mtcars_restored <- mtcars_merged |> 
  (\(df) df[names(mtcars)])() |>
  (\(df) { df[cols_to_num] <- lapply(df[cols_to_num], as.numeric); df })()
mtcars_restored
```

Confirm that the reconstructed data frame is identical to the original.

```{r}
identical(mtcars, mtcars_restored)
```

Confirm that the reconstructed data frame is identical to the original, ignoring the row order (a more relaxed equivalence test):

```{r}
nrow(setdiff(mtcars, mtcars_restored)) == 0
```

Create database.

```{r}
con <- dbConnect(RSQLite::SQLite(), "mtcars.sqlite")
```

```{r}
dbWriteTable(conn = con, name = "cars", value = cars_table)
dbWriteTable(con, "cyl", cyl_table)
dbWriteTable(con, "vs", vs_table)
dbWriteTable(con, "am", am_table)
dbWriteTable(con, "gear", gear_table)
dbWriteTable(con, "carb", carb_table)
dbWriteTable(con, "specs", specs_table)
```

```{r}
dbListTables(con)
```

Close the database connection.

```{r}
dbDisconnect(con)
```

Cleanup.

```{r}
file.remove("mtcars.sqlite")
```
