---
title: "Create database from data frame `mtcars`"
subtitle: "Using dplyr"
author: "Alex"
date: "`r format(Sys.time(), '%d %B %Y, %H:%M:%S')`"
output: 
  html_document:
    df_print: paged
    toc: true
    number_sections: false
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(RSQLite)
```

Print data frame overview.

```{r}
head(mtcars)
```

Load helper functions.

```{r}
source("helper.R")
```

Create augmented data frame.

```{r}
z <- mtcars %>%
  mutate(make = rownames(mtcars),
         id = as.integer(row_number()),
         cyl_id = as.intfac(cyl),
         vs_id = as.intfac(vs),
         am_id = as.intfac(am),
         gear_id = as.intfac(gear),
         carb_id = as.intfac(carb))
rownames(z) <- seq_len(nrow(z))
head(z)
```

Create table of cars specs (mpg, disp, hp, drat, wt, qsec).

```{r}
specs_table <- z %>%
  dplyr::select(id, mpg, disp:qsec)
head(specs_table)
```

Create table of cyl.

```{r}
cyl_table <- z %>%
  select(cyl, cyl_id) %>%
  distinct() %>%
  arrange(cyl) %>%
  mutate(cyl = as.integer(cyl))
cyl_table
```

Create table of vs.

```{r}
vs_table <- z %>%
  select(vs, vs_id) %>%
  distinct() %>%
  arrange(vs) %>%
  mutate(vs = as.integer(vs))
vs_table
```

Create table of am.

```{r}
am_table <- z %>%
  select(am, am_id) %>%
  distinct() %>%
  arrange(am) %>%
  mutate(am = as.integer(am))
am_table
```

Table of gear.

```{r}
gear_table <- z %>%
  select(gear, gear_id) %>%
  distinct() %>%
  arrange(gear) %>%
  mutate(gear = as.integer(gear))
gear_table
```

Table of carb.

```{r}
carb_table <- z %>%
  select(carb, carb_id) %>%
  distinct() %>%
  arrange(carb) %>%
  mutate(carb = as.integer(carb))
carb_table
```

Table of cars.

```{r}
cars_table <- z %>%
  select(id, make, cyl_id, vs_id, am_id, gear_id, carb_id) %>%
  distinct() %>%
  arrange(id)
head(cars_table)
```

Rejoin tables to recreate original data frame.

```{r}
mtcars_merged <- cars_table %>% 
  right_join(., cyl_table, by = "cyl_id") %>% 
  right_join(., vs_table, by = "vs_id") %>% 
  right_join(., am_table, by = "am_id") %>%
  right_join(., gear_table, by = "gear_id") %>% 
  right_join(., carb_table, by = "carb_id") %>% 
  select(c(id, make, cyl, vs, am, gear, carb)) %>% 
  right_join(., specs_table, by = "id") %>%
  as_tibble()
mtcars_merged <- as.data.frame(mtcars_merged)
head(mtcars_merged)
```

Restore row names.

```{r}
rownames(mtcars_merged) <- z$make
```

Reorder columns and fix data types.

```{r}
mtcars_restored <- mtcars_merged %>%
  select(all_of(names(mtcars))) %>%
  mutate(across(.cols = c(cyl, vs, am, gear, carb), .fns = ~ as.numeric(.x)))
head(mtcars_restored)
```

Confirm that the reconstructed data frame is identical to the original.

```{r}
identical(mtcars, mtcars_restored)
```

Confirm that the reconstructed data frame is identical to the original, ignoring the row order (a more relaxed equivalence test):

```{r}
nrow(setdiff(mtcars, mtcars_restored)) == 0
```

Create database.

```{r}
con <- dbConnect(RSQLite::SQLite(), "mtcars.sqlite")
```

```{r}
dbWriteTable(conn = con, name = "cars", value = cars_table)
dbWriteTable(con, "cyl", cyl_table)
dbWriteTable(con, "vs", vs_table)
dbWriteTable(con, "am", am_table)
dbWriteTable(con, "gear", gear_table)
dbWriteTable(con, "carb", carb_table)
dbWriteTable(con, "specs", specs_table)
```

```{r}
dbListTables(con)
```

Close the database connection.

```{r}
dbDisconnect(con)
```

Cleanup.

```{r}
file.remove("mtcars.sqlite")
```
